Finch: Sparse and Structured Array Programming with Control Flow

### 2024-04-02 sections/eval.tex
Update on Overleaf.

\subsubsection{Standardizing \& Heuristic Optimization}
Before a plan in Finch Logic can be interpreted, it must be converted to a standard form which resolves ambiguity about loop ordering and output formatting. The semantically important requirements of this standard form are 1) all inputs (i.e. tables and alias operators) in a query's RHS must conform to a common ordering of the indices 2) the outermost operator of each query's RHS must be a reformat 3) the expression within the reformat must be a pointwise expression, optionally wrapped in an aggregate operator. The former allows the interpreter to identify the loop order for each kernel. The second determines the output format for each intermediate. The last one guarantees that the innermost expression can be computed as a single kernel.

During the standardization process, a concordization pass is performed which examines each query, heuristically selects a loop order, and transposes inputs which don't match that order. Next, a formatting pass is performed which examines each query and selects a level format for each output index based on the formats of the inputs and whether random writes are required. The procedure for this is a simple set of rules which attempts to aggressively preserve the structure present in the input tensors.

>>>>

\subsubsection{Standardizing \& Heuristic Optimization}
Before a plan in Finch Logic can be interpreted, it must be converted to a standard form which resolves ambiguity about loop ordering and output formatting. The semantically important requirements of this standard form are 1) all inputs (i.e. tables and alias operators) in a query's RHS must conform to a common ordering of the indices 2) the outermost operator of each query's RHS must be a reformat 3) the expression within the reformat must be a pointwise expression, optionally wrapped in an aggregate operator. The former allows the interpreter to identify the loop order for each kernel. The second determines the output format for each intermediate. The last one guarantees that the innermost expression can be computed as a single kernel. To achieve this: 1) a concordization pass examines each query, heuristically selects a loop order, and transposes inputs which don't match that order 2) a formatting pass selects a level format for each output index based on the formats of the inputs and whether the loop order requires random writes. In the future,
